# Reproducible Research: Peer Assessment 1
=================================================

## Loading and preprocessing the data
```{r preprocess, echo = T}
unzip("activity.zip")
rawActivityData <- read.csv("activity.csv")
#rawActivityData$date <- as.POSIXlt(rawActivityData$date)
#for(i in 1:nrow(rawActivityData)) {
#  if(is.na(rawActivityData[i,1]) == T) {
#        rawActivityData[i,1] <- 0
#  }
#}
totalSteps <- tapply(rawActivityData$steps, rawActivityData$date, sum, na.rm = TRUE)
dfSteps <- as.data.frame(totalSteps)
dfSteps <- cbind(names(totalSteps), dfSteps)
names(dfSteps) <- c("Date", "Steps")
#observedDates  <- unique(rawActivityData$date[duplicated(rawActivityData$date)])
```
## What is mean total number of steps taken per day?
[https://class.coursera.org/repdata-004/forum/thread?thread_id=30#post-104](https://class.coursera.org/repdata-004/forum/thread?thread_id=30#post-104 "Clarification in the forums")
### 1. Make a histogram of the total number of steps taken each day
```{r histStepsEachDay, echo = TRUE}
hist(dfSteps$Steps,main="Histogram of Total Number of Steps taken per day", 
     xlab="Total Steps", col="orange")
```
### 2. Calculate and report the mean and median total number of steps taken per day
```{r meanMedianSteps, echo = T}
meanTotalNoOfStepsTakenPerDay <- mean(totalSteps, na.rm = T)
meanTotalNoOfStepsTakenPerDay
medianTotalNoOfStepsTakenPerDay <- median(totalSteps, na.rm = T)
medianTotalNoOfStepsTakenPerDay
```
The mean and median steps taken per day are `r meanTotalNoOfStepsTakenPerDay` and `r medianTotalNoOfStepsTakenPerDay` respectively.

## What is the average daily activity pattern?
### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r dailyActivityPattern , echo = T}
# calculate means per interval
totalStepsInterval <- tapply(rawActivityData$steps, rawActivityData$interval, mean, na.rm = T)

# create a data frame with the result
dfInterval <- as.data.frame(totalStepsInterval)
dfInterval <- cbind(names(totalStepsInterval), dfInterval)
names(dfInterval) <- c("Interval", "Steps")

## Create a column in the dfInterval to represent a Date Time for use in the time series plot
# pad intervals with leading 0's, so that each cell will consist of 4 numbers
require("stringr")
dfInterval$TimeInterval <- str_pad(dfInterval$Interval, width=4, pad="0")

# create datetime variable
require("lubridate")
dfInterval$TimeInterval <- parse_date_time(dfInterval$TimeInterval,"%H%M")

# Create a line plot with average daily activity
require("ggplot2")
require("scales")

p <- qplot(TimeInterval, Steps, data = dfInterval, geom = "line",   
           main="Daily Activity Pattern", xlab="Time slot") + 
           geom_line(colour="blue") + 
           scale_x_datetime(labels = date_format("%H:00"))
print(p)
``` 

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r maxSteps, echo = T}
maxSteps <- which.max(dfInterval$Steps)
interval <- dfInterval$Interval[maxSteps]
timeInterval <- format(dfInterval$TimeInterval[maxSteps],"%H:%M")
``` 

The interval with the most/maximum number of steps is `r interval`. 
The time interval corresponds to `r timeInterval`.
## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

### 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

## Are there differences in activity patterns between weekdays and weekends?
### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
